---
output:
  html_document: default
  pdf_document: default
---
```{r}
library(ggplot2)
```

Problem 7A
```{r}
for_problem_A <- c(1997, 906, 904, 32)
problem_A.matrix <- matrix(for_problem_A, nrow = 1, ncol = 4, byrow = TRUE)
chisq.test(problem_A.matrix, correct = FALSE)
```

Problem 7B
```{r}
males = c(3755, 3251, 3777, 3706, 3717, 3660, 3669, 3626, 3481, 3590, 3605, 3392)
exp_males = rep(3602.42, 12)
test_males = sum(((males - exp_males)**2)/exp_males)
test_males

females = c(1362, 1244, 1496, 1452, 1448, 1376, 1370, 1301, 1337, 1351, 1416, 1226)
exp_females = rep(1364.92, 12)
test_females = sum(((females - exp_females)**2)/exp_females)

qchisq(0.95, 10)
```

Problem 7C
```{r}
#a)
for_problem_C <- c(790 , 98 , 209 , 56 , 11 , 27, 21 , 7 , 12)
problem_c.matrix <- matrix(for_problem_C, nrow = 3, ncol = 3, byrow = TRUE)
chisq.test(problem_c.matrix, correct = FALSE)

#b)
chisq.test(problem_c.matrix, correct = FALSE)$expected
```

Problem 7D
```{r}
problem_7D_expected = chisq.test(problem_c.matrix, correct = FALSE)$expected
first_result = sum(((problem_c.matrix - problem_7D_expected)**2)/problem_7D_expected)
first_result
second_result = 2*sum(problem_c.matrix*log(problem_c.matrix/problem_7D_expected))
second_result
```

Problem 7H
```{r}
sample_size_1 = 1000
n=5
problem_H_data = rbinom(sample_size_1,n,prob=0.4)
problem_H_counts = (table(problem_H_data))
phat_MLE = sum(problem_H_data)/(sample_size_1*n)
exp_table = rep(1000*phat_MLE, 6)
chisq_calc = sum(((problem_H_counts - exp_table)**2)/exp_table)
1 - pchisq(chisq_calc, df = 5)
```

Problem 7I
```{r}
size_problem_I <- 2000
X_s = c()
Y_s = c()
for (i in 1:size_problem_I)
{
  ith_sample <- rbinom(sample_size_1,n,prob=0.4)
  ith_counts <- (table(ith_sample))
  ith_phat_MLE <- (sum(ith_counts)/(sample_size_1*n))
  ith_expected <- rep(1000*ith_phat_MLE, 6)
  X <- 2*sum(ith_counts*log(ith_counts/ith_expected))
  Y <- sum(((ith_counts - ith_expected)**2)/ith_expected)
  X_s <- append(X_s, X)
  Y_s <- append(Y_s, Y)
}

X_s <- data.frame(X_s)
Y_s <- data.frame(Y_s)

ggplot(X_s, aes(x=X_s)) + geom_histogram(aes(y=..density..), bins = 15) +
  stat_function(fun = dchisq, args = list(df = 5), col='blue')  

ggplot(Y_s, aes(x = Y_s)) + geom_histogram(aes(y=..density..), bins = 15) +
  stat_function(fun = dchisq, args = list(df = 5), col='blue') 
